Distributional Trees and Forests

### 2017-06-27 pkg/disttree/inst/useR2017/slides.Rnw
changed size of plots and formated density plot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1495 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber]{uibk}
\setfaculty{volkswirtschaftundstatistik}{en}

@@

\textbf{Distributional:} \\
\hspace{0.3cm} Specify the complete probability distribution\\
\hspace{0.3cm} (including location, scale and shape).

@@

\begin{frame}
\frametitle{Fitting a Tree}
\vspace*{-0.5cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
\vspace*{-0.21cm}
<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

@@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

@@

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.4, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.4, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dens_h, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE,
                          fill = "lightgray", col = "darkred", lwd = 1.5, ...) 
{
  ylines <- 3
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.5)
  xr <- xscale
  yr <- yscale

@@

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill))
    grid.lines(x = xd, y = yd, default.units = "native", 
               gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

@@

plot(tree, tnex = 1.4, drop = TRUE,
     terminal_panel = node_density(horizontal = TRUE, col = pal["forest"], lwd = 2),
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dens_v, fig=TRUE, echo=FALSE>>=
plot(tree, tnex = 1.4, drop = TRUE,
     terminal_panel = node_density(horizontal = FALSE, col = pal["forest"], lwd = 2),
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}
%\begin{frame}
%\frametitle{Fitting a Tree}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_box}
%\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "grey")
#plot(y=odata$y, x=odata$x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend(x = -0.0715, y = 21.65, expression(mu  %+-%  sigma), bty = "n")
#lines(x = odata$x, y = odata$mu + odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
#lines(x = odata$x, y = odata$mu - odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
@
\end{center}
\end{frame}

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}
%\setfaculty{volkswirtschaftundstatistik}{en}

@@

\textbf{Distributional:} \\
\vspace{0.1 cm}
\hspace{0.5cm} Specify the complete probability distribution\\
\hspace{0.5cm} (including location, scale and shape).

@@

\begin{frame}
\frametitle{Fitting a Tree}
\vspace*{-0.1cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
\vspace*{-0.21cm}
<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

@@

\setkeys{Gin}{width=0.65\linewidth, height=0.65\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.7)
@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.1cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.65\linewidth, height=0.65\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

@@

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.1cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.65\linewidth, height=0.65\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.7, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)], ylines = 1.5), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.1cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.65\linewidth, height=0.65\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.5)
  xr <- xscale
  yr <- yscale

@@

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

@@

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}
%\begin{frame}
%\frametitle{Fitting a Tree}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_box}
%\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-1.03cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "grey")
#plot(y=odata$y, x=odata$x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend(x = -0.0715, y = 21.62, expression(mu  %+-%  sigma), bty = "n")
#lines(x = odata$x, y = odata$mu + odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
#lines(x = odata$x, y = odata$mu - odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
@
\end{center}
\end{frame}
